---
title: "Fig1d_dnds_pancancer_pantissue"
output: html_document
date: "2024-02-13"
---


```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(ggrepel)
library(ggstatsplot)
library(dndscv)
library(ggpubr)
library(ggrepel)
library(readxl)
library(readr)
#Lucie
load("/Users/luciegourmet/Dropbox/Aim2_HallmarkTime/VAF_cancer_hallmarks/Relevant_analysis/Data/RefCDS_human_GRCh37_dNdScv.0.1.0.rda")
refdb_run =RefCDS
load("/Users/luciegourmet/Dropbox/Aim2_HallmarkTime/VAF_cancer_hallmarks/Relevant_analysis/Data/drivergenes.RData")
load("/Users/luciegourmet/Dropbox/Aim2_HallmarkTime/VAF_cancer_hallmarks/Relevant_analysis/Data/list_names.RData")
load("/Users/luciegourmet/Dropbox/Aim2_HallmarkTime/VAF_cancer_hallmarks/Relevant_analysis/Data/primary_data.RData")



```



```{r load prepare hallmarks list}

#stratify hallmark mutations
library(readxl)

Gene_list_of_cancer_hallmarks <- read_excel("/Users/luciegourmet/Dropbox/Aim2_HallmarkTime/VAF_cancer_hallmarks/Relevant_analysis/Data/Gene_list_of_cancer_hallmarks.xlsx")

#convert genes to ensembl id
library(biomaRt)
hsapiens_genes <- getBM(attributes = c("ensembl_gene_id", 
                                       "hgnc_symbol"),
                        mart = useMart("ensembl", dataset = "hsapiens_gene_ensembl"))

all_hallmarks <- hsapiens_genes %>%filter(hsapiens_genes$hgnc_symbol %in% Gene_list_of_cancer_hallmarks$ALL) 
all_hallmarks<-  unique(all_hallmarks$hgnc_symbol)
all_hallmarks<-as.data.frame(all_hallmarks)

angiogenesis <- hsapiens_genes %>%filter(hsapiens_genes$hgnc_symbol %in% Gene_list_of_cancer_hallmarks$ANGIOGENESIS) 
angiogenesis<-  unique(angiogenesis$hgnc_symbol)
angiogenesis<-as.data.frame(angiogenesis)
angiogenesis<-angiogenesis%>%filter(angiogenesis %in% list_names$list_names)


metabolism <- hsapiens_genes %>%filter(hsapiens_genes$hgnc_symbol %in% Gene_list_of_cancer_hallmarks$METABOLISM) 
metabolism<-  unique(metabolism$hgnc_symbol)
metabolism<-as.data.frame(metabolism)


death <- hsapiens_genes %>%filter(hsapiens_genes$hgnc_symbol %in% Gene_list_of_cancer_hallmarks$DEATH) 
death<-  unique(death$hgnc_symbol)
death<-as.data.frame(death)

proliferation <- hsapiens_genes %>%filter(hsapiens_genes$hgnc_symbol %in% Gene_list_of_cancer_hallmarks$PROLIFERATION) 
proliferation<-  unique(proliferation$hgnc_symbol)
proliferation<-as.data.frame(proliferation)


inflammation <- hsapiens_genes %>%filter(hsapiens_genes$hgnc_symbol %in% Gene_list_of_cancer_hallmarks$INFLAMMATION) 
inflammation<-  unique(inflammation$hgnc_symbol)
inflammation<-as.data.frame(inflammation)

metastasis <- hsapiens_genes %>%filter(hsapiens_genes$hgnc_symbol %in% Gene_list_of_cancer_hallmarks$METASTASIS) 
metastasis<-  unique(metastasis$hgnc_symbol)
metastasis<-as.data.frame(metastasis)

immortality <- hsapiens_genes %>%filter(hsapiens_genes$hgnc_symbol %in% Gene_list_of_cancer_hallmarks$IMMORTALITY) 
immortality<-  unique(immortality$hgnc_symbol)
immortality<-as.data.frame(immortality)


growth <- hsapiens_genes %>%filter(hsapiens_genes$hgnc_symbol %in% Gene_list_of_cancer_hallmarks$GROWTH) 
growth<-  unique(growth$hgnc_symbol)
growth<-as.data.frame(growth)


immune <- hsapiens_genes %>%filter(hsapiens_genes$hgnc_symbol %in% Gene_list_of_cancer_hallmarks$IMMUNE) 
immune<-  unique(immune$hgnc_symbol)
immune<-as.data.frame(immune)

mutations <- hsapiens_genes %>%filter(hsapiens_genes$hgnc_symbol %in% Gene_list_of_cancer_hallmarks$MUTATIONS) 
mutations<-  unique(mutations$hgnc_symbol)
mutations<-as.data.frame(mutations)


escape_genes = readxl::read_xlsx("/Users/luciegourmet/Dropbox/Projects/DriverLandscape/Data/SuppTable3_escapegenes_Aug2021.xlsx")
colnames(escape_genes)<-"escape"

#filter
death<-death%>%filter(death %in% list_names$list_names)
driver<-drivergenes%>%filter(drivergenes %in% list_names$list_names)
escape<-escape_genes%>%filter(escape %in% list_names$list_names)
growth<-growth%>%filter(growth %in% list_names$list_names)
immune<-immune%>%filter(immune %in% list_names$list_names)
immortality<-immortality%>%filter(immortality %in% list_names$list_names)
inflammation<-inflammation%>%filter(inflammation %in% list_names$list_names)
metabolism<-metabolism%>%filter(metabolism %in% list_names$list_names)
metastasis<-metastasis%>%filter(metastasis %in% list_names$list_names)
mutations<-mutations%>%filter(mutations %in% list_names$list_names)
proliferation<-proliferation%>%filter(proliferation %in% list_names$list_names)
angiogenesis<-angiogenesis%>%filter(angiogenesis %in% list_names$list_names)

```



```{r run dnds on each gene list of hallmarks}
#```{r load_functions}
#primary_data<-read_tsv("/Users/luciegourmet/Dropbox/Projects/DriverLandscape/TCGA_hg19_pancancer_annotated.tsv")
setwd("/Users/luciegourmet/Dropbox/Aim2_HallmarkTime/VAF_cancer_hallmarks/Relevant_analysis/Fig1b_dNdS_pancancer")
#primary_data$chr<-interaction( "chr", primary_data$chr, sep = "")
#create df



df_relevant_angiogenesis_dndscv<-dndscv(primary_data, refdb = refdb_run ,cv=NULL, gene_list =angiogenesis$angiogenesis)

save(df_relevant_angiogenesis_dndscv, file="df_relevant_angiogenesis_dndscv.RData")



df_relevant_metabolism_dndscv<-dndscv(primary_data, refdb = refdb_run ,cv=NULL, gene_list =metabolism$metabolism)

save(df_relevant_metabolism_dndscv, file="df_relevant_metabolism_dndscv.RData")



df_relevant_death_dndscv<-dndscv(primary_data, refdb = refdb_run ,cv=NULL, gene_list =death$death)

save(df_relevant_death_dndscv, file="df_relevant_death_dndscv.RData")



df_relevant_proliferation_dndscv<-dndscv(primary_data, refdb = refdb_run ,cv=NULL, gene_list =proliferation$proliferation)

save(df_relevant_proliferation_dndscv, file="df_relevant_proliferation_dndscv.RData")



df_relevant_inflammation_dndscv<-dndscv(primary_data, refdb = refdb_run ,cv=NULL, gene_list =inflammation$inflammation)

save(df_relevant_inflammation_dndscv, file="df_relevant_inflammation_dndscv.RData")



df_relevant_metastasis_dndscv<-dndscv(primary_data, refdb = refdb_run ,cv=NULL,gene_list =metastasis$metastasis)

save(df_relevant_metastasis_dndscv, file="df_relevant_metastasis_dndscv.RData")




df_relevant_immortality_dndscv<-dndscv(primary_data, refdb = refdb_run ,cv=NULL, gene_list =immortality$immortality)

save(df_relevant_immortality_dndscv, file="df_relevant_immortality_dndscv.RData")




df_relevant_growth_dndscv<-dndscv(primary_data, refdb = refdb_run ,cv=NULL, gene_list =growth$growth)

save(df_relevant_growth_dndscv, file="df_relevant_growth_dndscv.RData")



df_relevant_immune_dndscv<-dndscv(primary_data, refdb = refdb_run ,cv=NULL, gene_list =immune$immune)

save(df_relevant_immune_dndscv, file="df_relevant_immune_dndscv.RData")



df_relevant_mutations_dndscv<-dndscv(primary_data, refdb = refdb_run ,cv=NULL, gene_list =mutations$mutations)

save(df_relevant_mutations_dndscv, file="df_relevant_mutations_dndscv.RData")




df_relevant_driver_dndscv<-dndscv(primary_data, refdb = refdb_run ,cv=NULL, gene_list =driver$drivergenes)

save(df_relevant_driver_dndscv, file="df_relevant_driver_dndscv.RData")




df_relevant_escape_dndscv<-dndscv(primary_data, refdb = refdb_run ,cv=NULL, gene_list =escape$escape)

save(df_relevant_escape_dndscv, file="df_relevant_escape_dndscv.RData")

```



```{r plot the results for the 10 hallmarks}
plotdnds<-function(x){
  df_1<-x$globaldnds
  Globaldnds<- c(df_1[1,2])
  dndshigh<- c(df_1[1,4])
  dndslow<- c(df_1[1,3])
  df_x <- data.frame(Globaldnds, dndshigh, dndslow)
  return(df_x)
}

p_angiogenesis<-plotdnds(x =df_relevant_angiogenesis_dndscv)
p_angiogenesis$Hallmark<-"angiogenesis"

p_mutations<-plotdnds(x =df_relevant_mutations_dndscv)
p_mutations$Hallmark<-"genome instability"


p_immune<-plotdnds(x =df_relevant_immune_dndscv)
p_immune$Hallmark<-"immune evasion"

p_growth<-plotdnds(x =df_relevant_growth_dndscv)
p_growth$Hallmark<-"growth"

p_immortality<-plotdnds(x =df_relevant_immortality_dndscv)
p_immortality$Hallmark<-"immortality"


p_metastasis<-plotdnds(x =df_relevant_metastasis_dndscv)
p_metastasis$Hallmark<-"metastasis"

p_inflammation<-plotdnds(x =df_relevant_inflammation_dndscv)
p_inflammation$Hallmark<-"inflammation"

p_proliferation<-plotdnds(x =df_relevant_proliferation_dndscv)
p_proliferation$Hallmark<-"proliferation"

p_death<-plotdnds(x =df_relevant_death_dndscv)
p_death$Hallmark<-"death"

p_metabolism<-plotdnds(x =df_relevant_metabolism_dndscv)
p_metabolism$Hallmark<-"metabolism"

p_driver<-plotdnds(x =df_relevant_driver_dndscv)
p_driver$Hallmark<-"driver"

p_escape<-plotdnds(x =df_relevant_escape_dndscv)
p_escape$Hallmark<-"escape"


p_hallmarks<-rbind(p_metabolism, p_death, p_proliferation, p_inflammation, p_metastasis, p_immortality, p_growth, p_immune, p_mutations, p_angiogenesis)

Pancancer<-ggplot(p_hallmarks, aes(x= reorder(Hallmark, -Globaldnds), y = Globaldnds), color=Hallmark)+
  geom_errorbar(aes(ymin=dndslow, ymax=dndshigh, color=Hallmark), width=0, position=position_dodge(0.78))+ rotate_x_text(45) +   
  geom_point(position = position_dodge(0.78), size = 1.5) +  cowplot::theme_cowplot()+ geom_hline(yintercept=1, linetype="dashed", color = "red")+theme(legend.position="none")+scale_y_log10()+xlab(NULL) +ylab("dN/dS") +ggtitle("Cancer")+
  ggeasy::easy_center_title() + scale_color_manual(breaks = c("genome instability", "immortality", "metabolism", "angiogenesis","growth", "death","proliferation", "metastasis", "immune evasion", "inflammation"), values =c( "purple", "turquoise", "magenta", "red", "brown","blue","pink", "yellow", "orange", "green") )+
  theme(axis.text.x = element_text(angle=90, vjust=.5, hjust=1))

#+coord_flip()+ scale_x_discrete(limits = c("metastasis",  "proliferation", "inflammation", "death","immune evasion", "growth","angiogenesis", "metabolism", "genome instability","immortality"))

ggsave("Global_dnds_pancancer.pdf",plot = Pancancer, units = "cm",height = 12, width = 16)

```




---
```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(ggrepel)
library(ggstatsplot)
library(dndscv)
library(ggpubr)
library(ggrepel)
library(readxl)
library(readr)
#Lucie
load("/Users/luciegourmet/Dropbox/Aim2_HallmarkTime/VAF_cancer_hallmarks/Relevant_analysis/Data/RefCDS_human_GRCh37_dNdScv.0.1.0.rda")
refdb_run =RefCDS


```


```{r loadprepare}
load("/Users/luciegourmet/Dropbox/Aim2_HallmarkTime/VAF_cancer_hallmarks/Relevant_analysis/Data/drivergenes.RData")
load("/Users/luciegourmet/Dropbox/Aim2_HallmarkTime/VAF_cancer_hallmarks/Relevant_analysis/Data/list_names.RData")
load("/Users/luciegourmet/Dropbox/Aim2_HallmarkTime/VAF_cancer_hallmarks/Relevant_analysis/Data/annotated_normal_data.RData")
#stratify patients depending on whether they have hallmark mutations
library(readxl)

Gene_list_of_cancer_hallmarks <- read_excel("/Users/luciegourmet/Dropbox/Aim2_HallmarkTime/VAF_cancer_hallmarks/Relevant_analysis/Data/Gene_list_of_cancer_hallmarks.xlsx")

#convert genes to ensembl id
library(biomaRt)
hsapiens_genes <- getBM(attributes = c("ensembl_gene_id", 
                                       "hgnc_symbol"),
                        mart = useMart("ensembl", dataset = "hsapiens_gene_ensembl"))

all_hallmarks <- hsapiens_genes %>%filter(hsapiens_genes$hgnc_symbol %in% Gene_list_of_cancer_hallmarks$ALL) 
all_hallmarks<-  unique(all_hallmarks$hgnc_symbol)
all_hallmarks<-as.data.frame(all_hallmarks)

angiogenesis <- hsapiens_genes %>%filter(hsapiens_genes$hgnc_symbol %in% Gene_list_of_cancer_hallmarks$ANGIOGENESIS) 
angiogenesis<-  unique(angiogenesis$hgnc_symbol)
angiogenesis<-as.data.frame(angiogenesis)
angiogenesis<-angiogenesis%>%filter(angiogenesis %in% list_names$list_names)


metabolism <- hsapiens_genes %>%filter(hsapiens_genes$hgnc_symbol %in% Gene_list_of_cancer_hallmarks$METABOLISM) 
metabolism<-  unique(metabolism$hgnc_symbol)
metabolism<-as.data.frame(metabolism)


death <- hsapiens_genes %>%filter(hsapiens_genes$hgnc_symbol %in% Gene_list_of_cancer_hallmarks$DEATH) 
death<-  unique(death$hgnc_symbol)
death<-as.data.frame(death)

proliferation <- hsapiens_genes %>%filter(hsapiens_genes$hgnc_symbol %in% Gene_list_of_cancer_hallmarks$PROLIFERATION) 
proliferation<-  unique(proliferation$hgnc_symbol)
proliferation<-as.data.frame(proliferation)


inflammation <- hsapiens_genes %>%filter(hsapiens_genes$hgnc_symbol %in% Gene_list_of_cancer_hallmarks$INFLAMMATION) 
inflammation<-  unique(inflammation$hgnc_symbol)
inflammation<-as.data.frame(inflammation)

metastasis <- hsapiens_genes %>%filter(hsapiens_genes$hgnc_symbol %in% Gene_list_of_cancer_hallmarks$METASTASIS) 
metastasis<-  unique(metastasis$hgnc_symbol)
metastasis<-as.data.frame(metastasis)

immortality <- hsapiens_genes %>%filter(hsapiens_genes$hgnc_symbol %in% Gene_list_of_cancer_hallmarks$IMMORTALITY) 
immortality<-  unique(immortality$hgnc_symbol)
immortality<-as.data.frame(immortality)


growth <- hsapiens_genes %>%filter(hsapiens_genes$hgnc_symbol %in% Gene_list_of_cancer_hallmarks$GROWTH) 
growth<-  unique(growth$hgnc_symbol)
growth<-as.data.frame(growth)


immune <- hsapiens_genes %>%filter(hsapiens_genes$hgnc_symbol %in% Gene_list_of_cancer_hallmarks$IMMUNE) 
immune<-  unique(immune$hgnc_symbol)
immune<-as.data.frame(immune)

mutations <- hsapiens_genes %>%filter(hsapiens_genes$hgnc_symbol %in% Gene_list_of_cancer_hallmarks$MUTATIONS) 
mutations<-  unique(mutations$hgnc_symbol)
mutations<-as.data.frame(mutations)


escape_genes = readxl::read_xlsx("/Users/luciegourmet/Dropbox/Projects/DriverLandscape/Data/SuppTable3_escapegenes_Aug2021.xlsx")
colnames(escape_genes)<-"escape"

#filter
death<-death%>%filter(death %in% list_names$list_names)
driver<-drivergenes%>%filter(drivergenes %in% list_names$list_names)
escape<-escape_genes%>%filter(escape %in% list_names$list_names)
growth<-growth%>%filter(growth %in% list_names$list_names)
immune<-immune%>%filter(immune %in% list_names$list_names)
immortality<-immortality%>%filter(immortality %in% list_names$list_names)
inflammation<-inflammation%>%filter(inflammation %in% list_names$list_names)
metabolism<-metabolism%>%filter(metabolism %in% list_names$list_names)
metastasis<-metastasis%>%filter(metastasis %in% list_names$list_names)
mutations<-mutations%>%filter(mutations %in% list_names$list_names)
proliferation<-proliferation%>%filter(proliferation %in% list_names$list_names)
angiogenesis<-angiogenesis%>%filter(angiogenesis %in% list_names$list_names)

```




```{r run dnds on each gene list of hallmarks}

setwd("/Users/luciegourmet/Dropbox/Aim2_HallmarkTime/VAF_cancer_hallmarks/Relevant_analysis/Fig1b_dNdS_normal/dNdS_normal")




df_normal_angiogenesis_dndscv<-dndscv(annotated_normal_data, refdb = refdb_run ,cv=NULL, gene_list =angiogenesis$angiogenesis)

save(df_normal_angiogenesis_dndscv, file="df_normal_angiogenesis_dndscv.RData")



df_normal_metabolism_dndscv<-dndscv(annotated_normal_data, refdb = refdb_run ,cv=NULL, gene_list =metabolism$metabolism)

save(df_normal_metabolism_dndscv, file="df_normal_metabolism_dndscv.RData")



df_normal_death_dndscv<-dndscv(annotated_normal_data, refdb = refdb_run ,cv=NULL, gene_list =death$death)

save(df_normal_death_dndscv, file="df_normal_death_dndscv.RData")



df_normal_proliferation_dndscv<-dndscv(annotated_normal_data, refdb = refdb_run ,cv=NULL, gene_list =proliferation$proliferation)

save(df_normal_proliferation_dndscv, file="df_normal_proliferation_dndscv.RData")



df_normal_inflammation_dndscv<-dndscv(annotated_normal_data, refdb = refdb_run ,cv=NULL, gene_list =inflammation$inflammation)

save(df_normal_inflammation_dndscv, file="df_normal_inflammation_dndscv.RData")



df_normal_metastasis_dndscv<-dndscv(annotated_normal_data, refdb = refdb_run ,cv=NULL,gene_list =metastasis$metastasis)

save(df_normal_metastasis_dndscv, file="df_normal_metastasis_dndscv.RData")




df_normal_immortality_dndscv<-dndscv(annotated_normal_data, refdb = refdb_run ,cv=NULL, gene_list =immortality$immortality)

save(df_normal_immortality_dndscv, file="df_normal_immortality_dndscv.RData")




df_normal_growth_dndscv<-dndscv(annotated_normal_data, refdb = refdb_run ,cv=NULL, gene_list =growth$growth)

save(df_normal_growth_dndscv, file="df_normal_growth_dndscv.RData")



df_normal_immune_dndscv<-dndscv(annotated_normal_data, refdb = refdb_run ,cv=NULL, gene_list =immune$immune)

save(df_normal_immune_dndscv, file="df_normal_immune_dndscv.RData")



df_normal_mutations_dndscv<-dndscv(annotated_normal_data, refdb = refdb_run ,cv=NULL, gene_list =mutations$mutations)

save(df_normal_mutations_dndscv, file="df_normal_mutations_dndscv.RData")




df_normal_driver_dndscv<-dndscv(annotated_normal_data, refdb = refdb_run ,cv=NULL, gene_list =driver$drivergenes)

save(df_normal_driver_dndscv, file="df_normal_driver_dndscv.RData")




df_normal_escape_dndscv<-dndscv(annotated_normal_data, refdb = refdb_run ,cv=NULL, gene_list =escape$escape)

save(df_normal_escape_dndscv, file="df_normal_escape_dndscv.RData")

```



```{r plot the results for the 10 hallmarks}
plotdnds<-function(x){
  df_1<-x$globaldnds
  Globaldnds<- c(df_1[1,2])
  dndshigh<- c(df_1[1,4])
  dndslow<- c(df_1[1,3])
  df_x <- data.frame(Globaldnds, dndshigh, dndslow)
  return(df_x)
}
#get the mean and confidence interval for each hallmark
p_angiogenesis<-plotdnds(x =df_normal_angiogenesis_dndscv)
p_angiogenesis$Hallmark<-"angiogenesis"

p_mutations<-plotdnds(x =df_normal_mutations_dndscv)
p_mutations$Hallmark<-"genome instability"


p_immune<-plotdnds(x =df_normal_immune_dndscv)
p_immune$Hallmark<-"immune evasion"

p_growth<-plotdnds(x =df_normal_growth_dndscv)
p_growth$Hallmark<-"growth"

p_immortality<-plotdnds(x =df_normal_immortality_dndscv)
p_immortality$Hallmark<-"immortality"


p_metastasis<-plotdnds(x =df_normal_metastasis_dndscv)
p_metastasis$Hallmark<-"metastasis"

p_inflammation<-plotdnds(x =df_normal_inflammation_dndscv)
p_inflammation$Hallmark<-"inflammation"

p_proliferation<-plotdnds(x =df_normal_proliferation_dndscv)
p_proliferation$Hallmark<-"proliferation"

p_death<-plotdnds(x =df_normal_death_dndscv)
p_death$Hallmark<-"death"

p_metabolism<-plotdnds(x =df_normal_metabolism_dndscv)
p_metabolism$Hallmark<-"metabolism"

p_driver<-plotdnds(x =df_normal_driver_dndscv)
p_driver$Hallmark<-"driver"

p_escape<-plotdnds(x =df_normal_escape_dndscv)
p_escape$Hallmark<-"escape"

#only plot the 10 hallmarks
p_hallmarks<-rbind(p_metabolism, p_death, p_proliferation, p_inflammation, p_metastasis, p_immortality, p_growth, p_immune, p_mutations, p_angiogenesis)




Pancancer<-ggplot(p_hallmarks, aes(x= reorder(Hallmark, -Globaldnds), y = Globaldnds), color=Hallmark)+
  geom_errorbar(aes(ymin=dndslow, ymax=dndshigh, color=Hallmark), width=0, position=position_dodge(0.78))+ rotate_x_text(45) +   
  geom_point(position = position_dodge(0.78), size = 1.5) +  cowplot::theme_cowplot()+ geom_hline(yintercept=1, linetype="dashed", color = "red")+theme(legend.position="none")+scale_y_log10()+xlab("Cancer hallmarks (ordered by dN/dS)") +ylab("dN/dS") +ggtitle("Normal")+
  ggeasy::easy_center_title() + scale_color_manual(breaks = c("genome instability", "immortality", "metabolism", "angiogenesis","growth", "death","proliferation", "metastasis", "immune", "inflammation"), values =c("purple", "turquoise", "magenta", "red", "brown","blue","pink", "yellow", "orange", "green") )+
  theme(axis.text.x = element_text(angle=90, vjust=.5, hjust=1))+ scale_x_discrete(limits = c("immortality","genome instability","metabolism","angiogenesis","growth","immune evasion","death","inflammation","proliferation", "metastasis"))
#+coord_flip()+ scale_x_discrete(limits = c("metastasis","inflammation",  "immortality", "metabolism", "proliferation","growth", "death", "angiogenesis", "immune evasion",  "genome instability"))


ggsave("Global_dnds_normal.pdf",plot = Pancancer, units = "cm",height = 12, width = 16)

```
